Part III

### 2021-06-22 algebra.tex
added FHK stuff

\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}

@@

When we try to quantize, we run into an interesting phenomenon:
the Hamiltonian action $H_\rho^{\rm loc}$ does not extend as a Lie algebra map into the Weyl algebra $W_V$,
although it makes sense as a linear map.
The failure to respect the brackets determines, however, a cocycle on the loop algebra
by
\[
\alpha_V( x \otimes f(z) , y \otimes g(z)) = \Tr_V(xy) \Res(f \, \partial_z g).
\]
\owen{Check conventions.}
There is thus a central extension of the loop algebra 
\[
\CC c \to \widehat{\fg} \to \fg[z,z^{-1}]
\]
with a Lie algebra map
\[
\widehat{H}^{\rm loc}_\rho: \widehat{\fg} \to W_V,
\]
sometimes called {\em free field realization} of the affine Lie algebra~$\widehat{\fg}$.
This relationship --- and its extension to a map between the vertex algebras associated to $\widehat{\fg}$ and $V$ --- is a modest \owen{better adjective?} extract of the rich dialogue that has developed between representation theory, the physics of chiral conformal field theory, and the complex geometry of Riemann surfaces.

@@

>>>>

\def\U{{\rm U}}
\def\mc{\mathcal}
\def\mcol{\, | \,}
\def\ot{\otimes}
\def\Disj{\operatorname{Disj}}
\def\Open{\operatorname{Open}}
\def\Vect{\operatorname{Vect}}
\def\pAA{\mathring{\AA}}
\def\jou{{\mathtt{j}}}

@@

When we try to quantize, we run into an interesting phenomenon:
the Hamiltonian action $H_\rho^{\rm loc}$ does not extend as a Lie algebra map into the Weyl algebra $W_V$,
although it makes sense as a linear map.
The failure to respect the brackets determines, however, a cocycle on the loop algebra
by
\[
\alpha_V( x \otimes f(z) , y \otimes g(z)) = \Tr_V(xy) \Res(f \, \partial_z g).
\]
\owen{Check conventions.}
There is thus a central extension of the loop algebra 
\[
\CC c \to \widehat{\fg} \to \fg[z,z^{-1}]
\]
so that 
\[
[x \otimes f, y \otimes g] = [x,y] \otimes fg + \alpha_V( x \otimes f , y \otimes g) c,
\]
and there is a Lie algebra map
\[
\widehat{H}^{\rm loc}_\rho: \widehat{\fg} \to W_V,
\]
sometimes called {\em free field realization} of the affine Lie algebra~$\widehat{\fg}$.
This relationship --- and its extension to a map between the vertex algebras associated to $\widehat{\fg}$ and $V$ --- is a modest \owen{better adjective?} extract of the rich dialogue that has developed between complex analysis, representation theory, and the physics of chiral conformal field theory.
(To get Riemann surfaces into the game, we will need a richer framework than ordinary algebra.)

\owen{Discuss modules?}

@@

The story we have just told admits a natural extension to higher complex dimensions,
but it requires a first step in the direction of derived geometry.

Let $V$ denote a complex vector space (no longer required to be symplectic).
The naive idea is to replace the algebraic loop space $\Map(\CC^\times,  V)$ with a higher-dimensional analog $\Map(\CC^d -\{0\}, V)$.
If one considers ordinary algebraic or holomorphic maps for $d > 1$, 
then every such map extends across the origin, by Hartogs' lemma,
so this seems uninteresting.
One should notice, however, that punctured affine space $\mathring{\AA}^d = \CC^d -\{0\}$ is, as a scheme, not affine, 
and so the {\em derived} global sections of the structure sheaf $\cO_{\rm alg}$ are interesting:
\[
\RR \Gamma(\pAA^d, \cO_{\rm alg}) = H^* (\mathring{\AA}^d, \cO_{\rm alg}) 
= \begin{cases} 
0, & * \neq 0, d-1 \\ 
\CC[z_1,\ldots,z_d], & * = 0 \\ 
\CC[z_1^{-1},\ldots,z_d^{-1}] \frac{1}{z_1 \cdots z_d}, & * = d-1 
\end{cases},
\]
where we are providing a natural basis for the cohomology 
that aims to emphasize analogies with the $d=1$ case.
When $d=1$, note that this computation recovers the Laurent polynomials,
so when $d > 1$, 
we view the cohomology in degree $d-1$ as providing the derived replacement of the polar part of the Laurent polynomials.
Checking this computation is a straightforward exercise in algebraic geometry;
for instance, use the cover by the affine opens of the form $\AA^d \setminus \{z_i =0\}$.
\owen{I want to discuss the module structure on the $d-1$-component.}

A similar result holds in analytic geometry, of course,
so that we have a natural map
\[
\RR \Gamma(\pAA^d, \cO_{\rm alg}) \to \RR \Gamma(\pAA^d, \cO_{\rm an}) \simeq \Omega^{0,*}(\pAA^d)
\]
that replaces our inclusion of Laurent polynomials into the Dolbeault complex on~$\pAA^d$.

There is a natural analog of the residue map as well.
Just as $\CC^\times$ deformation-retracts onto the unit circle,
the punctured affine space $\pAA^d$ retracts onto the unit sphere $S^{2d-1}$.
The Bochner-Martinelli kernel 
\[
\omega_{BM} = \frac{(d-1)!}{(2 \pi i)^d} \frac{1}{(z\zbar)^d} \sum_{i=1}^d (-1)^{i-1} \zbar_i \d \zbar_1 \wedge \cdots \wedge \Hat{\d \zbar_i} \wedge \cdots \wedge \d \zbar_d
\]
determines a cohomology class of degree $d-1$, and we identify it with $1/z_1 \cdots z_d$ in our basis above,
so that $\omega_{BM}$ generates its cohomology group under multiplication by polynomials in the $z_j^{-1}$.
When $d = 1$, note that $\omega_{BM} = (1/2\pi i)\d z/z$, the kernel in the Cauchy residue formula.
It is also quick to check that
\[
\omega_{BM} \wedge \d z_1 \wedge \cdots \wedge \d z_d
\]
pulls back to a volume form on the unit sphere $S^{2d-1}$.
Hence, we define a residue map
\[
\Res: \RR \Gamma(\pAA^d, \cO_{\rm alg})  \to \CC
\]
by
\[
\Res(\alpha) = \oint_{S^{2d-1}} \alpha \wedge \d z_1 \wedge \cdots \wedge \d z_d.
\]
When $d =1$, this formula reduces to the usual residue map.
\owen{That's a bit misleading, maybe, but the integral really does only care about the cohomology class.}

The ingredients are thus in place to mimic the construction of the Weyl algebra for the formal loop space, of the affine Lie algebras, and of free field realization, but now with $\pAA^d$ in place of $\pAA^1 = \CC^\times$.
Reversing our treatment above, we first discuss the higher Kac-Moody algebras, 
following Faonte-Hennion-Kapranov \cite{FHK},
before discussing holomorphic field theory.

It is important to have an explicit dg commutative algebra that models the derived global sections,
and not just the cohomology groups.
There is a beautiful, explicit dg model $A_d$ for the algebraic version derived global sections due to Faonte-Hennion-Kapranov \cite{FHK} and based on the Jouanolou method for resolving singularities. 
In fact, they provide a model $A^{p,*}_d$ for the algebraic $p$-forms as well.
We will gloss their definition, referring the interested reader to their lucid paper for details.

Let $a_d$ denote the algebra  
\[
\CC[z_1,\ldots,z_d, z_1^*,\ldots,z_d^*][(z z^*)^{-1}]
\]
defined by localizing the polynomial algebra with respect to $zz^* = \sum_i z_i z^*_i$.
View this algebra $a_d$ as concentrated in bidegree $(0,0)$.
Consider the bigraded-commutative algebra $R^{*,*}_d$ over $a_d$ that is freely generated in bidegree $(1,0)$ by elements
\[
\d z_1,\ldots , \d z_d,
\] 
and in bidegree $(0,1)$ by
\[
\d z_1^*,\ldots, \d z_d^*.
\]
Note the close resemblance with the Dolbeault complex $\Omega^{*,*}(\CC^d)$ by thinking of $z_j^*$ as $\zbar_j$.
There is a subalgebra $A^{*,*}_d$ inside $\omega \in R^{p,m}_d$ and a natural pair of differentials $\partial$ and $\dbar$ that behave like the corresponding differentials in the Dolbeault complex.
We denote the subcomplex with $p=0$~by 
\[
(A_d, \dbar) = (\bigoplus_{q = 0}^d A_d^{q}[-q], \dbar),
\] 
and it has the structure of a dg commutative algebra.
For $p>0$, the complex $A^{p,*}_d = (\oplus_q A^{p,q}[-q], \dbar)$ is a dg module for $(A_d, \dbar)$.

The following proposition summarizes key properties of the dg algebra $A_d$ and its dg modules $A_{d}^{p,*}$,
by aggregating several results of~\cite{FHK}.

\begin{prop}[\cite{FHK}, Section 1]
\label{prop: Ad} $\;$
\begin{enumerate}
\item
The dg commutative algebra $(A_d,\dbar)$ is a model for $\RR \Gamma(\pAA^d, \sO_{\rm alg})$:
\[
A_d \simeq \RR\Gamma(\pAA^{d}, \sO^{alg}) .
\]
Similarly, $(A_d^{p,*},\dbar) \simeq \RR \Gamma(\pAA^{d}, \Omega^{p}_{\rm alg})$.
\item There is a dense map of commutative bigraded algebras
\[
\jou : A^{*,*}_d \to \Omega^{*,*}(\CC^d \setminus \{0\}),
\]
and the map intertwines with the $\dbar$ and $\partial$ differentials on both sides.
\item There is a unique $\GL_d$-equivariant residue map
\[
{\rm Res}_{z=0} : A_d^{d,d-1} \to \CC
\]
such that for any $\omega \in A^{d,d-1}_d$,
\[
{\rm Res}_{z=0} (\omega) = \oint_{S^{2d-1}} \jou(\omega)
\]
where $S^{2d-1}$ is any sphere centered at the origin in $\CC^d$. 
\item There is a representative $\omega_{BM}^{\rm alg}$ of the Bochner-Martinelli kernel such that
\[
\Res_{z=0} \left(f(z) \omega_{BM}^{alg}(z,z^*) \d z_1 \cdots \d z_d\right) = f(0)
\]
for any $f (z) \in \CC[z_1,\ldots,z_d]$. 
\end{enumerate}
\end{prop}

We can use this algebra and its nice properties to define a kind of ``sphere algebra'' $\Map(\pAA^d,\fg)$ of algebraic maps from the punctured affine space (or the punctured formal disk) into $\fg$,
just as $\fg[z,z^{-1}]$ provides a kind of loop algebra.

\begin{dfn}
For a Lie algebra $\fg$, the {\em sphere algebra} in complex dimension $d$ is the dg Lie algebra~$A_d \otimes \fg$.
Following \cite{FHK} we denote it by~$\fg^\bullet_d$.
\end{dfn}

There are natural central extensions of this sphere algebra as {\em $L_\infty$ algebras},
parallel to the affine Lie algebras appearing as central extensions of the loop algebra.
An $L_\infty$ algebra is a generalization of a Lie algebra;
it is graded vector space $L$ equipped with a countable collection of ``higher brackets'' $[-]_k: L^{\otimes k} \to L$ that satisfy a system of equations generalizing the Jacobi condition for a standard Lie algebra.
For an overview (and actual definition!) we recommend \owen{add citations}.

In our setting of $\fg^\bullet_d$, Faonte-Hennion-Kapranov introduced central extensions that only affect the $d$-bracket.
For any $\fg$-invariant degree $d+1$ polynomial  $\theta \in \Sym^{d+1}(\fg^*)^\fg$,  the FHK  cocycle is
\[
\label{fhk cocycle}
\begin{array}{cccc}
\theta_{\rm FHK} : & (A_d \tensor \fg)^{\tensor (d+1)} & \to & \CC\\ 
& a_0 \otimes \cdots \otimes a_d & \mapsto & \Res_{z=0} \theta(a_0,\partial a_1,\ldots,\partial a_d)
\end{array}.
\]
This cocycle has cohomological degree $2$ and so determines an unshifted central extension as $L_\infty$ algebras
\[
\CC c \to \widetilde{\fg}^\bullet_{d, \theta} \to \fg^\bullet_d ,
\]
so that the usual bracket (the 2-bracket) of $\fg^\bullet_d$ is unchanged but there is now a nontrivial $d$-bracket defined by~$\theta_{\rm FHK}$.
